---
title: "ST 2195 Qn 3"
output: html_document
---

# How does the number of people flying between different locations change over time?

## Import Packages
```{r}
install.packages("reshape2")
library("reshape2")
install.packages('ggplot2')
library("ggplot2")
install.packages("dplyr")
library(dplyr)
install.packages("tidyr")
library(tidyr)
```

## Importing Data
```{r}
table2006 <- read.csv("2006.csv") 
table2007 <- read.csv("2007.csv")
airports <- read.csv("airports.csv")
```

## Data Wrangling
We keep the columns that we will be using. we rename column 'iata' in airports data so that we can merge with the data
```{r}
df2006 <- table2006[,c("Year","Month","Origin","Dest")]
df2007 <- table2007[,c("Year","Month","Origin","Dest")]
dfairport <- airports[,c("iata","airport")]
dfairport1 <- dfairport %>% rename("Origin" = "iata")
airport2006 <- merge(df2006,dfairport1,by="Origin")
airport2007 <- merge(df2007,dfairport1,by="Origin")
```

## Busiest Airport
We will first find the airport with most departing flights. Then, we find the top 4 most traveled to airport from the airport with the most departing flights. 
```{r}
combinedairport <- rbind(airport2006,airport2007)

aggairport_tbl  <- combinedairport %>% group_by(Origin) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
dfairport_tbl <- aggairport_tbl %>% as.data.frame()
# We arrange the summarised table from above in descending order and slice only the first 3 rows.
dfbusiestairport <- dfairport_tbl %>% arrange(desc(total_count)) %>% slice(1:3)
dfbusiestairport
# Keep only the rows with 'ATL' in the column 'Origin'
dfATL <- combinedairport[combinedairport$Origin == 'ATL',]
# Repeat as done above to find the top 4 most traveled to airport.
aggdest_tbl  <- dfATL %>% group_by(Dest) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
dfdest_tbl <- aggdest_tbl %>% as.data.frame()
dfbusiestdest <- dfdest_tbl %>%                                     
  arrange(desc(total_count)) %>% slice(1:4)
dfbusiestdest  
```

## Data for Plotting
Now we use the information found above to get the data to plot the line graph. 
### 2006
```{r}
# Keep rows with value 'ATL' in the column 'Origin"
airport2006_ATL <- airport2006[airport2006$Origin == 'ATL',]
# Keep the rows with value 'LGA' in the column 'Dest' and the count of number of flights.
ATL_LGA2006 <- airport2006_ATL[airport2006_ATL$Dest == 'LGA',]
agg2006LGA_tbl  <- ATL_LGA2006 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
# Keep the rows with value 'DFW' in the column 'Dest'  and the count of number of flights.          
ATL_DFW2006 <- airport2006_ATL[airport2006_ATL$Dest == 'DFW',]
agg2006DFW_tbl  <- ATL_DFW2006 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
# Keep the rows with value 'MCO' in the column 'Dest' and the count of number of flights.
ATL_MCO2006 <- airport2006_ATL[airport2006_ATL$Dest == 'MCO',]
agg2006MCO_tbl  <- ATL_MCO2006 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
# Keep the rows with value 'EWR' in the column 'Dest' and the count of number of flights.
ATL_EWR2006 <- airport2006_ATL[airport2006_ATL$Dest == 'EWR',]
agg2006EWR_tbl  <- ATL_EWR2006 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
```

### 2007
Do the same as done in ### 2006 but with 2007 data.
```{r}
airport2007_ATL <- airport2007[airport2007$Origin == 'ATL',]

ATL_LGA2007 <- airport2007_ATL[airport2007_ATL$Dest == 'LGA',]
agg2007LGA_tbl  <- ATL_LGA2007 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
          
ATL_DFW2007 <- airport2007_ATL[airport2007_ATL$Dest == 'DFW',]
agg2007DFW_tbl  <- ATL_DFW2007 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')

ATL_MCO2007 <- airport2007_ATL[airport2007_ATL$Dest == 'MCO',]
agg2007MCO_tbl  <- ATL_MCO2007 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')

ATL_EWR2007 <- airport2007_ATL[airport2007_ATL$Dest == 'EWR',]
agg2007EWR_tbl  <- ATL_EWR2007 %>% group_by(Month) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
```

## Plotting
We will be plotting line graph with both years in the same graph for each airport.
```{r}
# Plotting the year 2006.
plot(agg2006LGA_tbl$Month, agg2006LGA_tbl$total_count, type='l', xlab='Month',ylab='Frequency',main='LGA', col = 'Blue',ylim=c(780,920))
# Plotting another in the same graph as above for year 2007.
lines(agg2007LGA_tbl$Month, agg2007LGA_tbl$total_count, type='l', col = 'Red')

plot(agg2006DFW_tbl$Month, agg2006DFW_tbl$total_count, type='l', xlab='Month',ylab='Frequency',main='DFW', col = 'Blue',ylim=c(680,900))
lines(agg2007DFW_tbl$Month, agg2007DFW_tbl$total_count, type='l', col = 'Red')

plot(agg2006MCO_tbl$Month, agg2006MCO_tbl$total_count, type='l', xlab='Month',ylab='Frequency',main='MCO', col = 'Blue',ylim=c(660,850))
lines(agg2007MCO_tbl$Month, agg2007MCO_tbl$total_count, type='l', col = 'Red')

plot(agg2006EWR_tbl$Month, agg2006EWR_tbl$total_count, type='l', xlab='Month',ylab='Frequency',main='EWR', col = 'Blue',ylim=c(530,750))
lines(agg2007EWR_tbl$Month, agg2007EWR_tbl$total_count, type='l', col = 'Red')
```
